Modelling the emergence of whisker barrels

### 2020-06-30 paper/briefpaper/brief.tex
Changes to param search paragraphs, figure and figure caption

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the model given
        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
        barrel area difference and $\omega$ (bottom row) is a measure of
        selectivity of the connection densities for the 41
        thalamocortical axon types. For each row, lighter (orange/yellow)
        colours imply a better model. The parameter space explored is three
        dimensional with the parameter ratio $\alpha/\beta$ varied along the
        columns of graphs, the competition parameter $\epsilon$ varied on the
        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
        asterisk (*) marks the parameter set displayed in
        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\paramsearch{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
  iterations plotted as colour maps for 216 parameter sets. Each row of maps
  shows the results for a separate metric. Each column of maps shows results
  for a different value of the ratio $\alpha / \beta$. The diffusion
  parameter, $D$, varies with the $x$ axes and the strength of competition,
  $\epsilon$ varies with the $y$ axes. The colour maps are chosen so that
  yellow indicates desirable properties for the simulations: Lower values of
  $\delta$ indicate good Voroni patterns and are yellow; Lower values of
  $\eta$ indicate good pattern matching; Higher values of $\omega$ indicate
  greater selectivity and are also coloured yellow. White squares indicate
  that the simulation was not stable for that parameter set.}

\paramsearch{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the
  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
  trade-off between good selection and good Voroni pattern
  conformation. The
  parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
  of varying $D$ in either direction from its optimum. ii)~is a less
  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

@@

\sensitivity{We performed a sensitivity analysis to determine if the quality of the
  pattern was affected by perturbations to any of the following inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
%
  In the following analysis, we always evaluated the three pattern metrics,
  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
  $\omega$ (which measures selectivity of connections, $c_i$) after running
  the model for 30000 simulation steps.}

@@

\cmnt{The pattern-matching metric, $\eta$, incorporates information about i)
  the differences in the areas of the simulated and experimental barrels
  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$); and ii)
  the information provided by `adjacency vectors', $\mathcal{V}_i$ (one for
  each barrel) whose $N$ elements contain the length of the barrel boundary
  which adjoins each other barrel in the pattern. The value of
  $\mathcal{V}_{i,j}$ (the $j^{\mathrm{th}}$ element of $\mathcal{V}_i$) is
  the length of the boundaries between regions of type $i$ and regions of type
  $j$. In a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector
  and note that $\mathcal{V}_{i,i} = 0, \forall i$. $\mathcal{V}_{i}$ can be
  computed for the experimental and simulated barrel patterns and compared. A
  dimensionless measure of the pattern's arrangement and neighbour relations
  can be obtained from the scalar product of the simulated and experimental
  adjacency vectors:
  $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where
  $b_i^{\mathrm{exp}}$($b_i^{\mathrm{sim}}$) is the total length of the boundary of
  experimental(simulated) barrel $i$. This measure is low early in the simulation, when the regions do
  not have the neighbour relations of the experimental pattern
  (Fig.\,\ref{fig:main}C, $t=1000$) and maximal when the neighbour relations
  have become established ($t=10000$). It does not distinguish between
  $t=10000$ and $t=30000$.}
%
\cmnt{A second way to analyse the adjacency vectors is to consider
  the mean magnitude of the difference between the simulated and experimental
  vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
    \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
  match and can separate patterns with straight boundaries from those with
  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B.ii).}
%
\cmnt{We combined all of these measures into a single pattern quality metric:}
%
\begin{equation} \label{eq:eta}
\color{blue}
  \eta =
  \frac
  {
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{A}_i^{\mathrm{sim}} - \mathcal{A}_i^{\mathrm{exp}} }
    \times
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{V}_{i}^{\mathrm{sim}} -
      \mathcal{V}_{i}^{\mathrm{exp}} }
  }
  {
    \frac{1}{N}
    \sum_i \frac {\mathcal{V}_{i}^{\mathrm{sim}}} {b_{i}^{\mathrm{sim}}}
    \vcdot
    \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
  },
\color{black}
\end{equation}
%
\cmnt{which has units of mm$^3$.}

>>>>

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the
        model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
        pattern metrics for $t=30000$ steps as colourmaps: Honda $\delta$ (top
        row) shows conformance to a Voronoi pattern; $\eta$ (middle row) gives
        the simulated/experimental barrel area difference and $\omega$ (bottom
        row) is a measure of selectivity of the connection densities for the
        41 thalamocortical axon types. The colour maps are chosen so that for
        each row, lighter (orange and yellow) colours indicate desirable
        properties for the simulations: Lower values of $\delta$ are yellow,
        indicating good Voroni patterns; Lower values of $\eta$ indicate good
        pattern matching; Higher values of $\omega$ indicate greater
        selectivity. White squares indicate that the simulation was not stable
        for that parameter set.
%
        The parameter space explored is three dimensional with the parameter
        ratio $\alpha/\beta$ varied along the columns of graphs, the
        competition parameter $\epsilon$ varied on the $y$-axes and the
        diffusion constant, $D$ varied along the $x$-axes. An asterisk (*)
        marks the parameter set displayed in Fig.\,\ref{fig:main}. Boxes i),
        ii)~and iii)~mark parameter sets for which corresponding patterns are
        shown in \textbf{B} (for $t=30000$ steps).
%
        \textbf{B} The effect of varying the diffusion constant,
        $D$. i)~Increasing $D$ above the optimum value gives an expansion of
        the outer barrels and a corresponding compression of the barrels
        within. ii)~Conversely, an expansion of the central barrels and
        compression of those near the boundary results from reducing $D$ from
        its optimum. iii) Reducing the diffusion constant is equivalent to
        increasing the size of the domain $S$ and so this simulation, in which
        $D$ had the lowest stable value, can be thought to represent an animal
        with a large cortical area but with unchanged cortical complexity. A
        large proportion of the domain is occupied by the projections with
        intermediate interaction parameters; those with strong interaction
        parameters are compressed around the edge of the domain. A barrel
        pattern fails to form.}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\paramsearch{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
  iterations, plotted as colour maps, for 216 parameter sets. The maps
  indicate that the pattern is not strongly influenced by the choice of
  $\alpha/\beta$, which was varied by 5 orders of magnitude. The most
  Voronoi-like patterns were achieved for parameter sets for which $D$ and
  $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the solution of
  maximum selectivity, which favours lower values of $D$. Thus, there is a
  trade-off between good selection and good Voroni pattern conformation.  The
  parameter set indicated by the asterisk in Fig.\,\ref{fig:paramsweep}A was
  chosen for Fig.\,\ref{fig:main} because it provides good pattern
  correspondence and a close approximation to a Voronoi tesselation.}

\paramsearch{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Because the patterns are robust
  to alterations in $\epsilon$ and the choices of $\alpha$ and $\beta$, we
  concentrate our attention on the diffusion constant, $D$. Changing $D$,
  whose value defines the rate at which a region of localized branching will
  spread out across the domain, $S$, is equivalent to making changes to the
  size of $S$ with a fixed branching diffusion rate. These simulations can
  therefore be thought of as representing animals with a varying cortical
  areas but with unchanged cortical complexity. The most extreme example
  (Fig.\,\ref{fig:paramsweep}B iii) could represent an animal with a very
  large cortical area (such as the South American capybara) which receives
  sensory input from $N=41$ thalamic barreloids. Here, a large proportion of
  the domain is occupied by the projections with intermediate interaction
  parameters and those with strong interaction parameters are bunched up
  towards the edge of the domain. In this case, a barrel pattern fails to
  form, agreeing with the observation that cortical whisker barrels are absent
  in the capybara.}

@@

% This section needs to emphasize that the absolute values of gamma are not important
\sensitivity{We performed a sensitivity analysis to determine if the quality of the
  pattern was affected by perturbations to any of the following inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
%
  In the following analysis, we always evaluated the three pattern metrics,
  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
  $\omega$ (which measures selectivity of connections, $c_i$) after running
  the model for 30000 simulation steps.}

@@

\metrics{The pattern-matching metric, $\eta$, incorporates information about i)
  the differences in the areas of the simulated and experimental barrels
  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$); and ii)
  the information provided by `adjacency vectors', $\mathcal{V}_i$ (one for
  each barrel) whose $N$ elements contain the length of the barrel boundary
  which adjoins each other barrel in the pattern. The value of
  $\mathcal{V}_{i,j}$ (the $j^{\mathrm{th}}$ element of $\mathcal{V}_i$) is
  the length of the boundaries between regions of type $i$ and regions of type
  $j$. In a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector
  and note that $\mathcal{V}_{i,i} = 0, \forall i$. $\mathcal{V}_{i}$ can be
  computed for the experimental and simulated barrel patterns and compared. A
  dimensionless measure of the pattern's arrangement and neighbour relations
  can be obtained from the scalar product of the simulated and experimental
  adjacency vectors:
  $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where
  $b_i^{\mathrm{exp}}$($b_i^{\mathrm{sim}}$) is the total length of the boundary of
  experimental(simulated) barrel $i$. This measure is low early in the simulation, when the regions do
  not have the neighbour relations of the experimental pattern
  (Fig.\,\ref{fig:main}C, $t=1000$) and maximal when the neighbour relations
  have become established ($t=10000$). It does not distinguish between
  $t=10000$ and $t=30000$.}
%
\metrics{A second way to analyse the adjacency vectors is to consider
  the mean magnitude of the difference between the simulated and experimental
  vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
    \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
  match and can separate patterns with straight boundaries from those with
  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B.ii).}
%
\metrics{We combined all of these measures into a single pattern quality metric:}
%
\begin{equation} \label{eq:eta}
\color{revred}
  \eta =
  \frac
  {
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{A}_i^{\mathrm{sim}} - \mathcal{A}_i^{\mathrm{exp}} }
    \times
    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{V}_{i}^{\mathrm{sim}} -
      \mathcal{V}_{i}^{\mathrm{exp}} }
  }
  {
    \frac{1}{N}
    \sum_i \frac {\mathcal{V}_{i}^{\mathrm{sim}}} {b_{i}^{\mathrm{sim}}}
    \vcdot
    \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
  },
\color{black}
\end{equation}
%
\metrics{which has units of mm$^3$.}
